---
title: "Visualization"
subtitle: "Part 2"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Remember this plot ..

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

## labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  )
```

## Scales

Start with some plot; x and y scales.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    trans = 'log'    #or trans = sqrt
  )
```

Look at color scales.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  scale_color_hue(
    name = "location",
    h = c(100,300)
  )
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE)
```

## Themes

Shift the legend.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

Change the overall theme.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE) +
  #theme_bw()         black and white background and grey grid lines
  #theme_minimal()    similar to the theme_bw, just get rid of the box around the ourside
  theme_classic() +
  theme(legend.position = 'bottom')  ## this should be after the theme_class() and the liked, or will be overrided
```

```{r}
weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature",
    y = "Maximun daily temperature",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "location",
    discrete = TRUE) +
  ggthemes::theme_excel()
```

## Setting options

This code truck should be palced in the beginning of each R markdown to set some default setting for the following contents.

```{r, eval=FALSE}  
library(tidyverse)

knitr::opt_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = '90%'
)

theme_set(theme_minimal() + theme(legend.posision = 'bottom'))

options(
  ggplot2.continous.colour = 'viridis_d',
  ggplot2.continous.fill = 'viridis_d'
)

scalr_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Data args in `geom`

```{r}
central_park = 
  weather_df %>% 
  filter(name == 'CentralPark_NY')

waikiki = 
  weather_df %>% 
  filter(name == 'Waikiki_HA')

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

## `patchwork`

Organize different plots into a single figure.

```{r}
tmax_tmin_df =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x =prcp, fill = name)) +
  geom_density(alpha = .5) +
  theme(legend.position = "none")

tmax_data_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none")

tmax_tmin_df + prcp_dens_p + tmax_data_p

## tmax_tmin_df + (prcp_dens_p + tmax_data_p)
## tmax_tmin_df / (prcp_dens_p + tmax_data_p)
## (tmax_tmin_df + prcp_dens_p) / tmax_data_p
```

# Data manipulation

Control your factos, and do the data manipulation to change the order the name

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c('Waikiki_HA'))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = .5)
```

What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = 'observation',
    values_to = 'temperatures'
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5)+
  facet_grid(. ~ name)
```

